Circumnuclear Media and Accetion Rates of Quiescent Supermassive Black Holes

### 2015-02-12 ms.tex
githole5

@@

Kinetic feedback occurs from an outflow of mass or energy from the
viscinity of the black hole in the form of a disk wind or jet, which
deposits its energy across a much wider range of radii (e.g.~\citealt{McNamara&Nulsen07}; \citealt{Novak+11}; \citealt{Gaspari+12}).  If the total
outflow power is proportional to the SMBH accretion rate, $L_{\rm j} =
\epsilon_{\rm j} \dot{M}c^{2}$, where $\epsilon_{\rm j} < 1$ is an
efficiency factor.  Assuming that this energy is deposited uniformly
out to a radius $r_{\rm heat}$ and volume $V_{\rm heat} \propto r_{\rm
heat}^{3}$, then the volumetric heating rate $e = \epsilon_{\rm
j}\dot{M}c^{2}/V_{\rm heat}$ results in an effective wind velocity at
the stagnation radius given by

\begin{eqnarray} v_{w}^{\bullet} &\approx& \left(\frac{2t_{\rm
h}\epsilon_{\rm j}\dot{M}c^{2}}{V_{\rm heat}\eta
\rho_{\star}}\right)^{1/2} \sim \left(\frac{\epsilon_{\rm
j}\dot{M}c^{2}t_{\rm h}}{\eta M_{\rm enc}}\right)^{1/2} \nonumber \\
&\approx& 300\,{\rm km\,s^{-1}}\,\left(\frac{\epsilon_{\rm j}}{10^{-6}}\right)^{1/2}\left(\frac{r_{\rm s}}{r_{\rm
heat}}\right)^{1-\Gamma/2},
\label{eq:vjet}
\end{eqnarray}

where we have used the fact that $\dot{M} \approx \eta M_{\rm
enc}(r_{\rm s})/t_{\rm h}$ and that $M_{\rm enc}(r) \propto
M_{\bullet}(r/r_{\rm soi})^{2-\Gamma}$ for radii interior to the outer
break radius (eq.~[\ref{eq:rhostar}]).  Thus for $r_{\rm heat} \sim
r_{\rm s}$, even a modest heating efficiency $\epsilon_{\rm j} \gtrsim
10^{-5}$ is sufficient for $v_{w}^{\bullet}$ to exceed other sources
of non-accretion powered heating.  However, if the jet distributes its
energy on much larger scales comparable to the size of the galaxy,
i.e. $r_{\rm heat} \gtrsim $ kpc $\gtrsim 10^{3}r_{\rm s}$ then the
heating rate would be much less significant, even for an efficient jet
($\epsilon_{\rm j} \sim 1$).

>>>>

\subsubsection{Compton Heating}

@@

\subsubsection{Kinetic Feedback}

Kinetic feedback results from outflows of energy or momentum from near the black hole in the form of a disk wind or jet, which deposits its energy across a potentially much larger range of radii (e.g.~\citealt{McNamara&Nulsen07}; \citealt{Novak+11}; \citealt{Gaspari+12}).

Assume that the jet or disk outflow power is proportional to the SMBH accretion rate, $L_{\rm j} =
\epsilon_{\rm j} \dot{M}c^{2}$, where $\epsilon_{\rm j} < 1$ is an efficiency factor.  If this energy is deposited as heat uniformly interior to a radius $r_{\rm heat}$ and volume $V_{\rm heat} \propto r_{\rm
heat}^{3}$, then the volumetric heating rate $e = \epsilon_{\rm
j}\dot{M}c^{2}/V_{\rm heat}$ results in an effective wind velocity at
the stagnation radius given by
\begin{eqnarray} v_{w}^{\bullet} &\approx& \left(\frac{2t_{\rm
h}\epsilon_{\rm j}\dot{M}c^{2}}{V_{\rm heat}\eta
\rho_{\star}}\right)^{1/2} \sim \left(\frac{\epsilon_{\rm
j}\dot{M}c^{2}t_{\rm h}}{\eta M_{\rm enc}}\right)^{1/2} \nonumber \\
&\approx& 300\,{\rm km\,s^{-1}}\,\left(\frac{\epsilon_{\rm j}}{10^{-6}}\right)^{1/2}\left(\frac{r_{\rm s}}{r_{\rm
heat}}\right)^{1-\Gamma/2},
\label{eq:vjet}
\end{eqnarray} 
where we have used the facts that $\dot{M} = \eta M_{\rm
enc}|_{r_{\rm s}}/t_{\rm h}$ and, for radii inside the Nuker break radius, $M_{\rm enc}(r) \propto
M_{\bullet}(r/r_{\rm soi})^{2-\Gamma}$ (eq.~[\ref{eq:rhostar}]).  If the outflow deposits much of its energy close to the sphere of influence, $r_{\rm heat} \sim
r_{\rm s}$, then even a small heating efficiency $\epsilon_{\rm j} \gtrsim
10^{-5}$ is sufficient for $v_{w}^{\bullet}$ to exceed other sources
of non-accretion powered heating.  However, if the disk outflow instead distributes energy over much larger scales comparable to the size of the galaxy, i.e. $r_{\rm heat} \gtrsim $ kpc $\sim 10^{3}r_{\rm s}$, then even for an efficient jet ($\epsilon_{\rm j} \sim 1$) the heating from kinetic feedback is unimportant compared to other local sources.
